### 2021-02-18 papers/thesis/full/ch2-spherical-harmonics/spherical-harmonics.tex
Edits to SH chapter

Jacobi operators that correspond to multiplication of the orthogonal polynomial basis, in this case the spherical harmonics, by our cartesian coordinates $x, y, z$. We start by expressing $x\,\Ylm(x,y,z)$, $y\,\Ylm(x,y,z)$ , and $z\,\Ylm(x,y,z)$ in terms of $Y^{m'}_{l'}(x,y,z)$ for any point $(x,y,z)$ on the unit sphere.

\begin{lemma}
	For $l \in \No$, $m \in \Z$ s.t. $0 \le |m| \le l$, the spherical harmonics satisfy the relationships:
\bseqnnumber{
	x\,\Ylm(x,y,z) &= \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) +  \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \nonumber \\
		     & \quad \quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) + \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z), \label{eqn:shrecx} \\
	y\,\Ylm(x,y,z) &= \beta_{l,m,1} Y^{m-1}_{l-1}(x,y,z) +  \beta_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \nonumber \\
		     & \quad \quad \quad + \beta_{l,m,3} Y^{m-1}_{l+1}(x,y,z) + \beta_{l,m,4} Y^{m+1}_{l+1}(x,y,z), \label{eqn:shrecy} \\
	z\,\Ylm(x,y,z) &= \gamma_{l,m,1} Y^{m}_{l-1}(x,y,z) + \gamma_{l,m,2} Y^{m}_{l+1}(x,y,z), \label{eqn:shrecz}
}
where
\bseqn{
	\alpha_{l,m,1} &:= 
		\begin{cases}
			\frac{\clm}{2c^{m-1}_{l-1}} \tilde \alpha_{l,m,1} \quad \text{if } m>0 \\
			\frac{\clm}{2c^{m-1}_{l-1}} \tilde \alpha_{l,|m|,2} \quad \text{if } m\le0 , \, l - |m| \ge 2 \\
			\quad \quad 0 \quad \quad \quad \quad \text{otherwise} 
		\end{cases} \\ 
	\alpha_{l,m,2} &:= 
		\begin{cases}
			\frac{\clm}{2c^{m+1}_{l-1}} \tilde \alpha_{l,m,2} \quad \text{if } m\ge0, \, l - |m| \ge 2 \\
			\frac{\clm}{2c^{m+1}_{l-1}} \tilde \alpha_{l,|m|,1} \quad \text{if } m<0 \\
			\quad \quad 0 \quad \quad \quad \quad \text{otherwise} 
		\end{cases} \\
	\alpha_{l,m,3} &:= 
		\begin{cases}
			\frac{\clm}{2c^{m-1}_{l+1}} \tilde \alpha_{l,m,3} \quad \text{if } m>0 \\
			\frac{\clm}{2c^{m-1}_{l+1}} \tilde \alpha_{l,|m|,4} \quad \text{if } m\le0 
		\end{cases} \\
	\alpha_{l,m,4} &:=
		\begin{cases}
			\frac{\clm}{2c^{m+1}_{l+1}} \tilde \alpha_{l,m,4} \quad \text{if } m\ge0 \\
			\frac{\clm}{2c^{m+1}_{l+1}} \tilde \alpha_{l,|m|,3} \quad \text{if } m<0 
		\end{cases} \\
	\beta_{l,m,j} &:= (-1)^{j+1} \: i \: \alpha_{l,m,j}, \quad j = 1,2,3,4 \\
	\gamma_{l,m,1} &:= \frac{\clm}{c^{m}_{l-1}} \: \tilde \gamma_{l,m,1} \\
	\gamma_{l,m,2} &:= \frac{\clm}{c^{m}_{l+1}} \: \tilde \gamma_{l,m,2}.
}
and
\bseqn{
	\tilde \alpha_{l,m,1} &:= \frac{2l}{2l+1} \\
	\tilde \alpha_{l,m,2} &:= 
		\begin{cases}
			- \frac{l}{2(2l+1)} \quad \text{if } l - m \ge 2 \\
			\quad \quad 0 \quad \quad \text{otherwise}
		\end{cases} \\
	\tilde \alpha_{l,m,3} &:= - \frac{2(l-m+2)(l-m+1)}{(2l+1)(l+1)} \\
	\tilde \alpha_{l,m,4} &:= \frac{(l+m+2)(l+m+1)}{2(2l+1)(l+1)} \\
	\tilde \gamma_{l,m,1} &:= \begin{cases}
			\frac{l}{2l+1} \quad \text{if } l - m \ge 1 \\
			\quad 0 \quad \quad \text{otherwise}
		\end{cases} \\
	\tilde \gamma_{l,m,2} &:= \frac{(l-m+1)(l+m+1)}{(2l+1)(l+1)}.
}
\end{lemma}

\begin{proof}
	The recurrence relationship for the Jacobi polynomials satisfies \cite[18.9.1, 18.9.2]{DLMF} 
\bseqn{
	z P^{(m,m)}_{l-m}(z) = \tilde \gamma_{l,m,1} P^{(m,m)}_{l-m-1}(z) + \tilde \gamma_{l,m,2} P^{(m,m)}_{l-m+1}(z),
}
for $k \ge 0, \, m \in \Z$. Hence, the expression for multiplication of the spherical harmonic $\Ylm$ by $z$ is
\bseqn{
	z\,\Ylmxyz &= \clm \eimtheta \rho(z)^{|m|} z P^{(|m|,|m|)}_{l-|m|}(z)  \\
		     &= \clm \eimtheta \rho(z)^{|m|} \big[ \tilde \gamma_{l,m,1} P^{(|m|,|m|)}_{l-|m|-1}(z) + \tilde \gamma_{l,m,2} P^{(|m|,|m|)}_{l-|m|+1}(z) \big] \\
		     &= \gamma_{l,m,1} Y^{m}_{l-1}(x,y,z) + \gamma_{l,m,2} Y^{m}_{l+1}(x,y,z).
}

For multiplication by $x$ and $y$, we require some further relations for the complex exponential and the Jacobi polynomials. First, recall that
\bseqn{
	\costheta \, \eimtheta &= \frac{1}{2} (e^{i\theta} + e^{-i\theta}) \eimtheta =  \frac{1}{2} (e^{i(m+1)\theta} + e^{i(m-1)\theta}) \\
	\sintheta \, \eimtheta &= \frac{1}{2i} (e^{i\theta} - e^{-i\theta}) \eimtheta =  \frac{-i}{2} (e^{i(m+1)\theta} - e^{i(m-1)\theta})
}
We can express a Jacobi polynomial $P^{(a,b)}_{l-m}(z)$ in terms of an Ultraspherical polynomial $C^{(\lambda)}_{l-m}(z)$ \cite[18.7.2]{DLMF}:
\bseqn{
	{l! \: (2m)! \over (l+m)! \: m!} \: C^{(m+1/2)}_{l-m}(z) = P^{(m,m)}_{l-m}(z),
}
where the Ultraspherical polynomials satisfy the relations \cite[18.9.7, 18.9.8]{DLMF}:
\bseqn{
	C^{(\lambda)}_n(z) &= {\lambda \over n + \lambda} \: \big[ C^{(\lambda+1)}_n(z) - C^{(\lambda+1)}_{n-2}(z) \big]  \\
	(1-z^2) \, C^{(\lambda)}_n(z) &= {1 \over 4(\lambda-1)(n+\lambda)} \: \big[ (n+2\lambda-2)(n+2\lambda-1) \: C^{(\lambda-1)}_n(z) \\
	&\quad \quad \quad \quad \quad \quad \quad \quad \quad \quad - (n+1)(n+2) \: C^{(\lambda-1)}_{n+2}(z) \big] ,
}
for $m \in \No$ where $\lambda > -\half$ is some parameter. Thus, combining these, we can write three-term recurrences for the Jacobi polynomials as:
\bseqn{
	P^{(m,m)}_{l-m}(z) &= \tilde \alpha_{l,m,4} P^{(m+1,m+1)}_{l-m}(z) + \tilde \alpha_{l,m,2} P^{(m+1,m+1)}_{l-m-2}(z) \\
	(1-z^2) \, P^{(m,m)}_{l-m}(z) &= \tilde \alpha_{l,m,3} P^{(m-1,m-1)}_{l-m+2}(z) + \tilde \alpha_{l,m,1} P^{(m-1,m-1)}_{l-m}(z),
}
for $l, m \in \No$ s.t. $0 \le m \le l$. Hence, the expressions for multiplication of the spherical harmonic $\Ylm$ by $x$ and $y$ are then:
\bseqn{
	x \,\Ylmxyz &= \clm \costheta \: \eimtheta \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= \half \clm (e^{i(m+1)\theta} + e^{i(m-1)\theta}) \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= \half \clm e^{i(m+1)\theta} \rho(z)^{|m|+1} \: \big[ \tilde \alpha_{l,m,4} P^{(|m|+1,|m|+1)}_{l-|m|}(z) + \tilde \alpha_{l,m,2} P^{(|m|+1,|m|+1)}_{l-|m|-2}(z) \big] \\
	&\quad \quad + \half \clm e^{i(m-1)\theta} \rho(z)^{|m|-1} \: \big[ \tilde \alpha_{l,m,3} P^{(|m|-1,|m|-1)}_{l-|m|+2}(z) + \tilde \alpha_{l,m,1} P^{(|m|-1,|m|-1)}_{l-|m|}(z) \big] \\
	&= \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) +  \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \\
	&\quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) + \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z), \\
	y \,\Ylmxyz &= \clm \sintheta \: \eimtheta \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= -\half i \: \clm (e^{i(m+1)\theta} - e^{i(m-1)\theta}) \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= i \: \Big\{ \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) - \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \\
	&\quad \quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) - \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z) \Big\}.
}

@@

We start as before for the scalar case by finding $x \,\gradYlm (x,y,z)$, $x \,\gradpYlm (x,y,z)$ etc. in terms of $\gradY^{m'}_{l'} (x,y,z)$, $\gradpY^{m'}_{l'} (x,y,z)$. A lovely consequence of our definitions of the scalar and vector spherical harmonics is that we can once again find explicit expressions for the coefficients in the aforementioned relations. To this end, we recall and derive some important relations of the complex exponential and Jacobi polynomials that are a part of the spherical harmonics definitions. In particular we first note that:
\bseqnnumber{
	\int_0^{2\pi} \: \eimtheta \: e^{-im'\theta} \: \cos(\theta) \D\theta &= \pi (\delta_{m',m-1} + \delta_{m',m+1}) \label{eqn:intcomplexexpcos} \\
	\int_0^{2\pi} \: \eimtheta \: e^{-im'\theta} \: \sin(\theta) \D\theta &= i \; \pi (\delta_{m',m-1} - \delta_{m',m+1}) \label{eqn:intcomplexexpsin}
}
and
\bseqnnumber{
	&\int_{-1}^1 \: \jaclmabs{l}{m}(z) \: \jaclmabs{l'}{m}(z) \: \rho(z)^{2\abs{m}} \: \D z = \delta_{l',l} \: {1 \over 2\pi (\clm)^2} \label{eqn:normjac} \\
	&\jaclmabsdz{l}{m}(z) = d_{l,m} \: \jaclmbracket{l}{\abs{m} + 1} \quad \text{where } d_{l,m} := \half (l + \abs{m} + 1), \label{eqn:diffjac},
}
where \bsrefeqn{eqn:normjac} and \bsrefeqn{eqn:normjac} are consequences of the definition of the Jacobi polynomials, see \cite[Table 18.3.1, 18.9.15]{DLMF}. One final relation we will need is that for incrementing the Jacobi parameters, which we will present in the following Lemma.
\begin{lemma}
	Let $\alpha, \beta \in \R$ be general parameters. Let $z \in [-1,1]$ and define $P_{-2}^{(\alpha+1, \beta+1)}(z) \equiv P_{-1}^{(\alpha+1, \beta+1)}(z) \equiv 0$. Then, for any $n \in \N_0$:
\bseqn{
	P_{n}^{(\alpha, \beta)}(z) =
		{1 \over 2n + \alpha + \beta + 1} \Big\{ & {(n+\alpha+\beta+1)(n+\alpha+\beta+2) \over 2n+\alpha+\beta+2} \: P_{n}^{(\alpha+1, \beta+1)}(z) \\
									  & + \big[{(n+\alpha+\beta+1)(n+\alpha) \over 2n+\alpha+\beta} \\
									  		& \quad \quad - {(n+\alpha+\beta+1)(n+\beta+1) \over 2n+\alpha+\beta+2}\big] \: P_{n-1}^{(\alpha+1, \beta+1)}(z) \\
									  & + {(n+\alpha)(n+\beta) \over 2n+\alpha+\beta} \: P_{n-2}^{(\alpha+1, \beta+1)}(z) \Big\}.
}
\end{lemma}
\begin{proof}
	This is simply as consequence of the relationships established for incrementing the parameters of Jacobi polynomials detailed in \cite[18.9.5]{DLMF}, along with the symmetry relation that $P_{n}^{(\alpha, \beta)}(-z) \equiv (-1)^n P_{n}^{(\beta, \alpha)}(z)$.
\end{proof}
We thus can present a useful corollary specific for our needs.
\begin{corollary}\label{corollary:incrementjacparams}
	Let $l, m \in \N_0$ s.t. $0 \le m \le l$ and let $z \in [-1,1]$. Then:
\bseqn{
	\jaclm{l}{m}(z) = \mu_{l,m,1} \: P_{l-m}^{(m+1, m+1)}(z) + \mu_{l,m,2} \: P_{l-m-2}^{(m+1, m+1)}(z),
}
where
\bseqn{
	\mu_{l,m,1} &:= {(l+m+1)(l+m+2) \over 2(2l + 1)(l+1)} \\
	\mu_{l,m,2} &:= {l \over 2(2l + 1)}.
}
\end{corollary}

>>>>

Jacobi operators that correspond to multiplication of the orthogonal polynomial basis, in this case the spherical harmonics, by our cartesian coordinates $x, y, z$. We can obtain the entries to these operator matrices by determining the coefficients of $x\,\Ylm(x,y,z)$, $y\,\Ylm(x,y,z)$, and $z\,\Ylm(x,y,z)$ in terms of $Y^{m'}_{l'}(x,y,z)$ for any point $(x,y,z)$ on the unit sphere.

First though, we require a few results concerning the incrementing and decrementing of the parameters for the Jacobi polynomials, and the three-term recurrence for the Jacobi polynomials, that will also prove useful later on in this chapter.
\begin{lemma}
	Let $\alpha, \beta \in \R$ be general parameters. Let $z \in [-1,1]$ and define $P_{-2}^{(\alpha+1, \beta+1)}(z) \equiv P_{-1}^{(\alpha+1, \beta+1)}(z) \equiv 0$. Then, for any $n \in \N_0$:
\bseqn{
	P_{n}^{(\alpha, \beta)}(z) =
		{1 \over 2n + \alpha + \beta + 1} \Big\{ & {(n+\alpha+\beta+1)(n+\alpha+\beta+2) \over 2n+\alpha+\beta+2} \: P_{n}^{(\alpha+1, \beta+1)}(z) \\
									  & + \big[{(n+\alpha+\beta+1)(n+\alpha) \over 2n+\alpha+\beta} \\
									  		& \quad \quad - {(n+\alpha+\beta+1)(n+\beta+1) \over 2n+\alpha+\beta+2}\big] \: P_{n-1}^{(\alpha+1, \beta+1)}(z) \\
									  & + {(n+\alpha)(n+\beta) \over 2n+\alpha+\beta} \: P_{n-2}^{(\alpha+1, \beta+1)}(z) \Big\}, \\
	P_{n}^{(\alpha, \beta)}(z) =
		{1 \over 2n + \alpha + \beta + 1} \Big\{ & {(n+\alpha+\beta+1)(n+\alpha+\beta+2) \over 2n+\alpha+\beta+2} \: P_{n}^{(\alpha+1, \beta+1)}(z) \\
									  & + \big[{(n+\alpha+\beta+1)(n+\alpha) \over 2n+\alpha+\beta} \\
									  		& \quad \quad - {(n+\alpha+\beta+1)(n+\beta+1) \over 2n+\alpha+\beta+2}\big] \: P_{n-1}^{(\alpha+1, \beta+1)}(z) \\
									  & + {(n+\alpha)(n+\beta) \over 2n+\alpha+\beta} \: P_{n-2}^{(\alpha+1, \beta+1)}(z) \Big\}, \\
	z \, P_{n}^{(\alpha, \beta)}(z) = 
		\Big\{ & {2(n+\alpha)(n+\beta) \over (2n + \alpha + \beta)(2n + \alpha + \beta + 1)} \: P_{n-1}^{(\alpha, \beta)}(z) \\
			 & {(\beta^2 - \alpha^2) \over (2n + \alpha + \beta)(2n + \alpha + \beta + 2)} \: P_{n}^{(\alpha, \beta)}(z) \\
			 & {2(n+1)(n + \alpha + \beta + 1) \over 2(2n + \alpha + \beta + 1)(2n + \alpha + \beta + 2)} \: P_{n+1}^{(\alpha, \beta)}(z).
}
\end{lemma}
\begin{proof}
	\bstodoinline{Do the second result (i.e. $(1-z^2)P_n$ relationship).}
	The first two results are simply consequences of the relationships established for incrementing and decrementing the parameters of Jacobi polynomials detailed in \cite[18.9.5, 18.9.6]{DLMF}, along with the symmetry relation that $P_{n}^{(\alpha, \beta)}(-z) \equiv (-1)^n P_{n}^{(\beta, \alpha)}(z)$. The final one is simply a rearranging of the established three-term recurrence detailed in \cite[18.9.1, 18.9.2]{DLMF}
\end{proof}
We thus can present a useful corollary specific for our needs.
\begin{corollary}\label{corollary:incrementjacparams}
	Let $l, m \in \N_0$ s.t. $0 \le m \le l$ and let $z \in [-1,1]$. Then:
\bseqn{
	(1-z^2) \, \jaclm{l}{m}(z) &=  \tilde \alpha_{l,m,1} P^{(m-1,m-1)}_{l-m}(z) + \tilde \alpha_{l,m,3} P^{(m-1,m-1)}_{l-m+2}(z) \\
	\jaclm{l}{m}(z) &= \tilde \alpha_{l,m,2} P^{(m+1,m+1)}_{l-m-2}(z) + \tilde \alpha_{l,m,4} P^{(m+1,m+1)}_{l-m}(z), \\
	z \, \jaclm{l}{m}(z) &= \tilde \gamma_{l,m,1} P^{(m,m)}_{l-m-1}(z) + \tilde \gamma_{l,m,2} P^{(m,m)}_{l-m+1}(z),
}
where
\bseqn{
	\tilde \alpha_{l,m,1} &:= {2l \over 2l+1} \\
	\tilde \alpha_{l,m,2} &:= 
		\begin{cases}
			- {l \over 2(2l+1)} \quad \text{if } l - m \ge 2 \\
			\quad \quad 0 \quad \quad \text{otherwise}
		\end{cases} \\
	\tilde \alpha_{l,m,3} &:= - {2(l-m+2)(l-m+1) \over (2l+1)(l+1)} \\
	\tilde \alpha_{l,m,4} &:= {(l+m+2)(l+m+1) \over 2(2l+1)(l+1)} \\
	\tilde \gamma_{l,m,1} &:= \begin{cases}
			\frac{l}{2l+1} \quad \text{if } l - m \ge 1 \\
			\quad 0 \quad \quad \text{otherwise}
		\end{cases} \\
	\tilde \gamma_{l,m,2} &:= \frac{(l-m+1)(l+m+1)}{(2l+1)(l+1)}.
}
\end{corollary}

\bsrefcor{corollary:incrementjacparams} allows us to easily determine the aforementioned coefficients, that we will formalise in the following Lemma.

\begin{lemma}\label{lemma:shxyzrelations}
	For $l \in \No$, $m \in \Z$ s.t. $0 \le |m| \le l$, the spherical harmonics satisfy the relationships:
\bseqnnumber{
	x\,\Ylm(x,y,z) &= \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) +  \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \nonumber \\
		     & \quad \quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) + \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z), \label{eqn:shrecx} \\
	y\,\Ylm(x,y,z) &= \beta_{l,m,1} Y^{m-1}_{l-1}(x,y,z) +  \beta_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \nonumber \\
		     & \quad \quad \quad + \beta_{l,m,3} Y^{m-1}_{l+1}(x,y,z) + \beta_{l,m,4} Y^{m+1}_{l+1}(x,y,z), \label{eqn:shrecy} \\
	z\,\Ylm(x,y,z) &= \gamma_{l,m,1} Y^{m}_{l-1}(x,y,z) + \gamma_{l,m,2} Y^{m}_{l+1}(x,y,z), \label{eqn:shrecz}
}
where
\bseqn{
	\alpha_{l,m,1} &:= 
		\begin{cases}
			\frac{\clm}{2c^{m-1}_{l-1}} \tilde \alpha_{l,m,1} \quad \text{if } m>0 \\
			\frac{\clm}{2c^{m-1}_{l-1}} \tilde \alpha_{l,|m|,2} \quad \text{if } m\le0 , \, l - |m| \ge 2 \\
			\quad \quad 0 \quad \quad \quad \quad \text{otherwise} 
		\end{cases} \\ 
	\alpha_{l,m,2} &:= 
		\begin{cases}
			\frac{\clm}{2c^{m+1}_{l-1}} \tilde \alpha_{l,m,2} \quad \text{if } m\ge0, \, l - |m| \ge 2 \\
			\frac{\clm}{2c^{m+1}_{l-1}} \tilde \alpha_{l,|m|,1} \quad \text{if } m<0 \\
			\quad \quad 0 \quad \quad \quad \quad \text{otherwise} 
		\end{cases} \\
	\alpha_{l,m,3} &:= 
		\begin{cases}
			\frac{\clm}{2c^{m-1}_{l+1}} \tilde \alpha_{l,m,3} \quad \text{if } m>0 \\
			\frac{\clm}{2c^{m-1}_{l+1}} \tilde \alpha_{l,|m|,4} \quad \text{if } m\le0 
		\end{cases} \\
	\alpha_{l,m,4} &:=
		\begin{cases}
			\frac{\clm}{2c^{m+1}_{l+1}} \tilde \alpha_{l,m,4} \quad \text{if } m\ge0 \\
			\frac{\clm}{2c^{m+1}_{l+1}} \tilde \alpha_{l,|m|,3} \quad \text{if } m<0 
		\end{cases} \\
	\beta_{l,m,j} &:= (-1)^{j+1} \: i \: \alpha_{l,m,j}, \quad j = 1,2,3,4 \\
	\gamma_{l,m,1} &:= \frac{\clm}{c^{m}_{l-1}} \: \tilde \gamma_{l,m,1} \\
	\gamma_{l,m,2} &:= \frac{\clm}{c^{m}_{l+1}} \: \tilde \gamma_{l,m,2}.
}
\end{lemma}

\begin{proof}
	The expression for multiplication of the spherical harmonic $\Ylm$ by $z$ is simply a consequence of the third result in \bsrefcor{corollary:incrementjacparams}:
\bseqn{
	z\,\Ylmxyz &= \clm \eimtheta \rho(z)^{|m|} z P^{(|m|,|m|)}_{l-|m|}(z)  \\
		     &= \clm \eimtheta \rho(z)^{|m|} \big[ \tilde \gamma_{l,m,1} P^{(|m|,|m|)}_{l-|m|-1}(z) + \tilde \gamma_{l,m,2} P^{(|m|,|m|)}_{l-|m|+1}(z) \big] \\
		     &= \gamma_{l,m,1} Y^{m}_{l-1}(x,y,z) + \gamma_{l,m,2} Y^{m}_{l+1}(x,y,z).
}

For multiplication by $x$ and $y$, note that the complex exponential satisfies
\bseqn{
	\costheta \, \eimtheta &= \frac{1}{2} (e^{i\theta} + e^{-i\theta}) \eimtheta =  \frac{1}{2} (e^{i(m+1)\theta} + e^{i(m-1)\theta}) \\
	\sintheta \, \eimtheta &= \frac{1}{2i} (e^{i\theta} - e^{-i\theta}) \eimtheta =  \frac{-i}{2} (e^{i(m+1)\theta} - e^{i(m-1)\theta})
}
Combining this with the results in \bsrefcor{corollary:incrementjacparams}, the expressions for multiplication of the spherical harmonic $\Ylm$ by $x$ and $y$ are then:
\bseqn{
	x \,\Ylmxyz &= \clm \costheta \: \eimtheta \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= \half \clm (e^{i(m+1)\theta} + e^{i(m-1)\theta}) \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= \half \clm e^{i(m+1)\theta} \rho(z)^{|m|+1} \: \big[ \tilde \alpha_{l,m,4} P^{(|m|+1,|m|+1)}_{l-|m|}(z) + \tilde \alpha_{l,m,2} P^{(|m|+1,|m|+1)}_{l-|m|-2}(z) \big] \\
	&\quad \quad + \half \clm e^{i(m-1)\theta} \rho(z)^{|m|-1} \: \big[ \tilde \alpha_{l,m,3} P^{(|m|-1,|m|-1)}_{l-|m|+2}(z) + \tilde \alpha_{l,m,1} P^{(|m|-1,|m|-1)}_{l-|m|}(z) \big] \\
	&= \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) +  \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \\
	&\quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) + \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z),
}
and
\bseqn{
	y \,\Ylmxyz &= \clm \sintheta \: \eimtheta \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= -\half i \: \clm (e^{i(m+1)\theta} - e^{i(m-1)\theta}) \: \rho(z)^{|m|+1} P^{(|m|,|m|)}_{l-|m|}(z) \\
	&= i \: \Big\{ \alpha_{l,m,1} Y^{m-1}_{l-1}(x,y,z) - \alpha_{l,m,2} Y^{m+1}_{l-1}(x,y,z) \\
	&\quad \quad \quad + \alpha_{l,m,3} Y^{m-1}_{l+1}(x,y,z) - \alpha_{l,m,4} Y^{m+1}_{l+1}(x,y,z) \Big\}.
}
\end{proof}

@@

We start as before for the scalar case by finding $x \,\gradYlm (x,y,z)$, $x \,\gradpYlm (x,y,z)$ etc. in terms of $\gradY^{m'}_{l'} (x,y,z)$, $\gradpY^{m'}_{l'} (x,y,z)$. A lovely consequence of our definitions of the scalar and vector spherical harmonics is that we can once again find explicit expressions for the coefficients in the aforementioned relations. To this end, we recall and derive some important relations of the complex exponential and Jacobi polynomials that are a part of the spherical harmonics definitions. In particular we first note that:
\bseqnnumber{
	\int_0^{2\pi} \: \eimtheta \: e^{-im'\theta} \: \cos(\theta) \D\theta &= \pi (\delta_{m',m-1} + \delta_{m',m+1}) \label{eqn:intcomplexexpcos} \\
	\int_0^{2\pi} \: \eimtheta \: e^{-im'\theta} \: \sin(\theta) \D\theta &= i \; \pi (\delta_{m',m-1} - \delta_{m',m+1}) \label{eqn:intcomplexexpsin}
}
and
\bseqnnumber{
	&\int_{-1}^1 \: \jaclmabs{l}{m}(z) \: \jaclmabs{l'}{m}(z) \: \rho(z)^{2\abs{m}} \: \D z = \delta_{l',l} \: {1 \over 2\pi (\clm)^2} \label{eqn:normjac} \\
	&\jaclmabsdz{l}{m}(z) = d_{l,m} \: \jaclmbracket{l}{\abs{m} + 1} \quad \text{where } d_{l,m} := \half (l + \abs{m} + 1), \label{eqn:diffjac},
}
where \bsrefeqn{eqn:normjac} and \bsrefeqn{eqn:diffjac} are consequences of the definition of the Jacobi polynomials, see \cite[Table 18.3.1, 18.9.15]{DLMF}.

The expressions for the VSHs as given in \bsrefeqn{eqn:VSHs} together with the above allow us to be able to find the coefficients for our desired expressions for $x \gradYlm$, $y \gradYlm$, $z \gradYlm$, $x \gradpYlm$, $y \gradpYlm$ and $z \gradpYlm$.
